### 2019-04-07 Sections/intro_to_QIT.tex
+quantum internet project sec.
initi save

\comment{To do}

>>>>

\famousquote{When you find yourself in a room surrounded by your enemies you tell yourself, `I am not locked in here with you, you are locked in here with me'. This is the kind of mindset you should have if you want to succeed in life. Get rid of that victim mentality.}{Bruce Lee}

% \paragraph{Classical correlation measures}
\subsection{Probability and information}

Suppose we have a random variable $X$, and a particular measurement outcome $x$ occurs with probability $p(x)$, the Shannon entropy of $X$ is given by,
\begin{align}\index{Shannon entropy}
H(X) = -\sum_x p_x\log_2(p_x).
\end{align}

The information, or entropy, plays a central role in information theory. The intuitive interpretation is that it quantifies an experimenter's uncertainty about $X$ before measuring it, and his expected information gain is $H(X)$ bits upon learning the outcome. The information $H(X)$ is zero if and only if one of the probabilities $p(x)$ is unity, with the others being zero. In this case the value of $X$ is already known and so there is no information to be gained from observing it.

 For a joint distribution over $X$ and $Y$ this generalises to,
\begin{align}\index{Joint Shannon entropy}
H(X,Y) =  -\sum_{x,y} p_{x,y}\log_2(p_{x,y}).
\end{align}

The von Neuman entropy\index{von Neuman entropy} \cite{bib:bengtsson2017geometry} for quantum density operators, $S(\hat\rho)$, is defined analogously, replacing probabilities with density operator eigenvalues,
\begin{align}\index{von Neuman entropy}
S(\hat\rho) &= - \sum_x \lambda_x \log_2 (\lambda_x) \nonumber \\
&= -\mathrm{tr}(\hat\rho\,\log \,\hat\rho),
\end{align}
where $\{\lambda\}$ is the eigenvalue spectrum of $\hat\rho$. This modification is logically justified, as the eigenvalues can be interpreted directly as a purely classical probability distribution of orthogonal states when the density operator is transformed into a basis with no coherences between basis states (i.e a diagonal basis or spectral decomposition\index{Spectral decomposition}). In that case, the Shannon and von Neuman entropies essentially have identical physical interpretations.

We now introduce an entropic measure of common or mutual information shared by two parties. Suppose Alice possesses the random variable $A$ and Bob has the random variable $B$, the \textit{mutual information} specifies the number of bits in common between the two distributions. Equivalently, this represents the maximum number of bits that one party can learn about the other just by inspecting their own information.

For two classical distributions, the classical mutual information is given by,
\begin{align}\index{Mutual information}
I(A;B) = H(A) + H(B) - H(A,B).
\end{align}

It is a measure of the correlation between the events A and B. If these correspond, respectively, to the selection and receipt of a signal then H(A : B) is the information transferred by the communication.

Analogously, the quantum mutual information\index{Quantum mutual information} for bipartite state $\hat\rho_{A,B}$ is defined as,
\begin{align}
I(A;B)_{\hat\rho} = S(\hat\rho_A) + S(\hat\rho_B) - S(\hat\rho_{A,B}),
\end{align}
using the von Neuman entropy.

The mutual information between two quantum states is invariant under local unitary transformations,
\begin{align}
I(A;B)_{\hat\rho} = I(\hat{U}_A\hat\rho_A \hat{U}_A^\dag; \hat{U}_B\hat\rho_B \hat{U}_B^\dag),
\end{align}
since the eigenvalue spectrum of a density operator is invariant under unitary transformations. Therefore, the mutual information represents the maximum amount of information Bob can learn about Alice's state under \textit{any} local operations.

A quantum process cannot increase the mutual information between two parties. This yields the \textit{data processing inequality}\index{Data processing inequality} that, for a sequence of channels \mbox{$X\to Y\to Z$},
\begin{align}\index{Data processing inequality}\label{eq:data_proc_ineq}
I(X:Z)&\leq I(X:Y), \nonumber \\
I(X:Z)&\leq I(Y:Z),
\end{align}
with equality if and only if the channel not specified in the identity on the right hand side (\mbox{$Y\to Z$} or \mbox{$X\to Y$} respectively) is unitary, i.e one of the links in the chain perfectly preserves information content. The progression is shown in Fig.~\ref{fig:data_proc_ineq}.

\begin{figure}[!htbp]
\includegraphics[clip=true, width=0.3\textwidth]{data_proc_ineq}
\captionspacefig \caption{\label{fig:data_proc_ineq}A sequence of events \mbox{$X\to Y\to Z$}. The data processing inequality states that the mutual information from beginning to end is upper-bounded by the mutual information between neighbouring stages, as per Eq.~(\ref{eq:data_proc_ineq}).}	
\end{figure}

The mutual information is defined as being between a particular known pair of states. Of course, in a quantum channel we have the flexibility to manipulate the input state and the measurement at the output. From this, we can then define the \textit{classical information of the channel}\index{Classical information of the channel}, $I_c(\hat\rho_{A,B})$ -- the maximum of the mutual information, optimised over \textit{all} possible measurement settings. Suppose $\hat\rho_{A,B}$ is shared between Alice and Bob, from which they would like to extract maximal correlations (i.e joint information). They measure using local measurement bases $\{\Lambda_A^x\}$ and $\{\Lambda_B^x\}$, yielding random variables $A$ and $B$. Their quantum mutual information of the channel is defined as,
\begin{align}
I_c(\hat\rho_{A,B}) = \max_{\rho,\{\Lambda_A^x\},\{\Lambda_B^x\}} I(A;B). \label{eq:quant_mut}
\end{align}

The intuitive interpretation is that this is the maximum mutual information between input and output states that can be achieved across the channel. This effectively places a physical upper bound on the achievable bitrate or bandwidth of the channel.
